Switching Optimisation in Huffman Code for Power Efficient Data Transmission

### 2015-02-26 Switche Paper 1st approache/elsarticle-template-num.tex
some work

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{algcompatible}
\usepackage{multirow}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

@@

The rest of the paper is organised as follows:
Section 2 presents the background study of the
relationship of switching activity with power consumption, the relation of switching with Huffman code and the issues in biological data transmission. The proposed approach is described in Section 3. Experimental results and
discussion are presented in Section 4 . Finally,
concluding remarks are presented in Section 5. 
\section{Background}
\subsection{Effect of Switching in Power Consumption}
\subsection{Huffman Code and Its Relationship with Switching Activity}
%\subsubsection{Relationship between switching and Huffman Code}

>>>>

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{algcompatible}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{algcompatible}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

@@

The rest of the paper is organised as follows:
Section 2 presents the background study of the
relationship of switching activity with power consumption, the relation of switching with Huffman code and the issues in biological data transmission. The proposed approach is described in Section 3. Experimental results and
discussion are presented in Section 4 . Finally,
concluding remarks are presented in Section 5. 
\section{Background}
\subsection{Effect of Switching in Power Consumption}
CMOS is the most widely used technology implemented in VLSI chips because of their power efficiency. Power dissipation in CMOS circuits has three different components: dynamic, static, and leakage \cite{Weste88}.
\begin{equation*}
P_{avg} = P_{dynamic}+P_{static}+P_{leakge} ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\end{equation*}
\begin{equation}
~~~~~~~~~~~~= 0.5 \times C_{L} \times V_{dd}^{2}\times E(sw)\times f_{clk} + I_{sc} \times V_{dd}+I_{leakage} \times V_{dd}
 \end{equation}
 where\\
  $C_{L}$ is the load capacitance, \\
  $V_{dd}$ is power supply,\\
  $E(sw)$ is average transition (switching) number per clock cycle ,\\
  $f_{clk}$ is the clock frequency,\\
  $ I_{sc}$ is the short circuit current, and\\
  $ I_{leakage}$ is the leakage current.\\
Out of the three components, the dynamic power dissipation is the dominant and it counts for 70\%-90\% of the total power consumption\cite{pedram1996}. Therefore, optimisation of dynamic power consumption is considered in this paper. In CMOS circuit, dynamic power consumption arises when the load capacitance ($C_{L}$) is charged to address a transition from low to the high voltage level.  The parameters $C_{L}$, $V_{dd}$, and $f_{clk}$ in dynamic power dissipation are primarily determined by the circuit layout and the fabrication technology. However, amount of switching, $E(sw)$ completely depends on the digital representation of the data. As the dynamic power dissipation is proportional  to the average number of switching, thus a reduction in the total switching number will eventually results in the reduction of the dynamic power consumption. Therefore, in this paper, we focus on reducing total amount of switching in Huffman code to improve power efficiency of digital data processing.
%In this paper, by switching we mean a transition from `0' to `1' or `1' to `0' in a binary sequence. 
\subsection{Huffman Code and Its Relationship with Switching Activity}
%\subsubsection{Relationship between switching and Huffman Code}
In computer science and information theory, Huffman code is an entropy encoding algorithm used for lossless data compression. The classical Huffman code requires to pass the input data two times to complete the encoding operation. In the first pass, it reads the whole stream to determine the occurrence frequencies of distinct symbols. After that, symbols along with their frequencies are used to generate a binary tree known as Huffman tree. The process of Huffman tree generation is shown in algorithm 1 \cite{Cormen2001}.

\begin{algorithm}[!thpb]
\label{alg1}
\caption{Huffman(C)}
\begin{algorithmic}[1]
\State $n \leftarrow |C|$
\State $Q \leftarrow C$
\FOR {$i = 1~to~n-1$}
\State $allocate~a~new~node~z$
\State $left[z]$ $\leftarrow$ $x$ $\leftarrow$ $EXTRACT\_MIN (Q)$
\State $right[z]$ $\leftarrow$ $y$ $\leftarrow$ $EXTRACT\_MIN (Q)$
\State $f[z]$ $\leftarrow$ $f[x] + f[y]$
\State $INSERT (Q~,~z)$
\ENDFOR
\State \textbf{$retur$n} $EXTRACT-MIN (Q)$
\end{algorithmic}
\end{algorithm}
In the algorithm, C is the set of n distinct symbols, and each symbol $ c\in C$ has a frequency denoted as f[c]. The algorithm starts its operation with a set of $|C|$  leaf nodes where each node corresponds to a distinct symbol. After that, a sequence of $|C|-1$ merging operation is performed to obtain the final tree. The merging is done by extracting two least frequent symbols from the minimum priority queue, $Q$, and creating and inserting a new node (parent) to the queue for that two symbols. The frequency of the new  node is calculated as the sum of the frequency of its child nodes.  The resultant tree ensures the optimal encoding of symbols to obtain maximal compression performance. The tree is traversed from the root to the leaf nodes to obtain the unique codeword (binary string) for different symbols. Once the codewords are obtained, in the second pass, the Huffman algorithm encodes the symbols of the input message with the codewords, thus reduce the size of the whole message.
For a set of symbols more than one Huffman tree is possible, i.e., optimal solution can be obtained from different structure of the tree. If the structure of the tree is changed, the codewords for the symbols will also change. The switching activity depends on the structure of the codewords. For example, if a symbol is assigned a codewrod `000' then there is no switching, on the other hand, if the codeword `010' is assigned to a symbol the transmission of this symbol requires 2 switching activity.
